Classless Java

### 2015-11-18 paper/sections/Formalization.tex
polishing formalization

\marco{Old notes: new text is later}
Yanlin and Haoyuan

We need to show 2 things:

1) The dynamic semantics: what's the code that gets generated by a mixin annotation;

2) The type system: what programs to reject; properties: generation of type-safe/checkable code.

\bruno{The implementation is still missing the type system (rejecting some
  programs)!}

\bruno{Some notes for discussion regarding formalization:
\begin{itemize}
\item Syntax and Rules:
We can make it more compact and consistent by use overline
notation more, instead of \ldots.
\item Order of premisses: It may be worthwhile to swap the order of
some premisses, so that it follows the order of the explanations.
\item T-Invk: I think we need to type-check $e_0$ separately. Otherwise
$C_0$ shows up magically.
\item T-Obj: small fixes on 3rd line. Suggestion, create two
auxiliary functions: $sigvalid(C,\overline{mh})$ and $alldefined(C,\overline{mh})$
\item T-INTF: $C_0$ is $I$. Type of this was $C$, modified it to $C_0$.
In the last line, should it be $dom(\overline{meth})$ or just $\overline{meth}$?

@@

\item I have trouble following the definition of shadow; I think there
are errors there.

@@

\begin{figure}[h]
\begin{grammar}
\production{
\e
}{
  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\x\QM=\e\QM;\e'\mid\obj
  }{expressions}\\
\production{
\obj
}{
\QM{new}\ \C\oR\cR\oC\fields\
\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
\ldots
\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
\cC
  }{object creation}\\
\production{\field}{\T\ \f \QM= \x\QM;}{field declaration}\\
\production{
\II
}{
 \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\ \methods\ \cC
  }{interface declaration}\\
\production{
\method
}{
 \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
\mid
\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
\mid
\mh\QM{;}
  }{method declaration}\\
\production{
\mh
}{
 \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
  }{method header}\\
\production{
\ann
}{
  \mixinAnn|\emptyset
  }{annotations}\\
\production{\Gamma}{
\x_1{:}\C_1\ldots\x_n{:}\C_n
}{environment}
\end{grammar}
\caption{Grammar of ClassLess Java}
\label{Grammar}
\end{figure}

${}_{}$\\
\begin{figure}[h]
$
\!\!\!\!\!\!\!\!\!\!
\begin{array}{l}
\inferrule[(T-Invk)]{
 \Gamma \vdash \e \in \C \\\\
\forall i\in 1..n\  \Gamma \vdash \e_i \in \_<:\C_i \\\\
  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
%\textsf{mmodifier}(\m,\C) \neq \textbf{static}
 }{
 \Gamma \vdash \e\QM.\m\QM(\overline{\e}\QM) \in \C' }
\quad
\inferrule[(T-StaticInvk)]{
\forall i\in 1..n\  \Gamma \vdash \e_i\in \_<:\C_i \\\\
\textsf{mtypeS}(\m,\C) \!=\! \overline{\C} \!\to\! \C'
%\textsf{mmodifier}(\m,\C) = \textbf{static}
}{
\Gamma \vdash \C\QM.\m\QM(\overline{\e}\QM) \in \C'}
\quad
\inferrule[(T-SuperInvk)]{
\Gamma(\this) <: \C \\\\
\forall i\in 1..n\ \Gamma \vdash \e_i\in \_<:\C_i \\\\
  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
%\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
}{\Gamma \vdash \C\QM.\QM{super}\QM.\m\QM(\overline{\e}\QM) \in \C'}

\\[5ex]
\inferrule[(T-Obj)]{
\forall i\in 1..n\
%\Gamma_i
\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
\vdash\e_i\in \_\subtype\C^{\mh_i}\\\\
\forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
\forall i\in 1..k\ \Gamma(\x_i)\subtype\T_i\\\\
\forall\m\mbox{ such that }
\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
%\forall i\in 1\ldots n\ \Gamma_i=\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
}{
\Gamma \vdash\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM=\x_1\QM;\ldots\T_k\ \f_k\QM=\x_k\QM;\
\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
\ldots
\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
\cC
\in\C
}
\quad
\inferrule[(T-Var)]{
\Gamma(\x)=\C
}{
\Gamma \vdash \x \in\C}
\\[5ex]

\quad
\inferrule[(T-update)]{
\Gamma \vdash \e\in\_<:\Gamma(\x)\\\\
\Gamma \vdash \e'\in\C
}{
\Gamma \vdash \x\QM=\e\QM;\e'\in\C }
\quad
 \inferrule[(T-Intf)]{
IT(\C_0) = \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n \oC\methods\cC\\\\
 \forall \QM{default}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
\ \Gamma^{\mh},\,\QM{this}{:}\C_0\vdash\e\in \_\subtype\C^{\mh} \\\\
 \forall \QM{static}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
\ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)
 }{
\C_0 \text{ OK}
}
\end{array}$
\caption{Typing}
\label{ET}
\end{figure}
\subsection{Grammar and typing rules}${}_{}$\\

>>>>

\bruno{Some notes for discussion regarding formalization:
\begin{itemize}
\item Order of premisses: It may be worthwhile to swap the order of
some premisses, so that it follows the order of the explanations.
\item T-Obj: small fixes on 3rd line. Suggestion, create two
auxiliary functions: $sigvalid(C,\overline{mh})$ and $alldefined(C,\overline{mh})$
\item T-INTF: $C_0$ is $I$. Type of this was $C$, modified it to $C_0$.
In the last line, should it be $dom(\overline{meth})$ or just $\overline{meth}$?

@@

@@

\begin{figure}[h]
\begin{grammar}
\production{
\e
}{
  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\x\QM=\e\QM;\e'\mid\obj
  }{expressions}\\
\production{
\obj
}{
\QM{new}\ \C\oR\cR\oC\fields\
\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
\ldots
\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
\cC
  }{object creation}\\
\production{\field}{\T\ \f \QM= \x\QM;}{field declaration}\\
\production{
\II
}{
 \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC
  }{interface declaration}\\
\production{
\method
}{
 \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
\mid
\QM{default}\ \mh\ \oC\QM{return}\ \e\QM{;}\cC
\mid
\mh\QM{;}
  }{method declaration}\\
\production{
\mh
}{
 \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
  }{method header}\\
\production{
\ann
}{
  \mixinAnn|\emptyset
  }{annotations}\\
\production{\Gamma}{
\x_1{:}\C_1\ldots\x_n{:}\C_n
}{environment}
\end{grammar}
\caption{Grammar of ClassLess Java}
\label{Grammar}
\end{figure}

\bruno{Use space better in this figure. }

${}_{}$\\
\begin{figure}[h]
$
\!\!\!\!\!\!\!\!\!\!
\begin{array}{l}

%% T-Invk
\inferrule[(T-Invk)]{
 \Gamma \vdash \e \in \C \\\\
\forall i\in 1..n\\   \Gamma \vdash \e_i \in \_<:\C_i \\\\
  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
%\textsf{mmodifier}(\m,\C) \neq \textbf{static}
 }{
 \Gamma \vdash \e\QM.\m\QM(\overline{\e}\QM) \in \C' }
\quad\quad

%%T-StaticInvk
\inferrule[(T-StaticInvk)]{
\forall i\in 1..n\  \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
\textsf{mtypeS}(\m,\C) \!=\! \overline{\C} \!\to\! \C'
%\textsf{mmodifier}(\m,\C) = \textbf{static}
}{
\Gamma \vdash \C\QM.\m\QM(\overline{\e}\QM) \in \C'}
\quad\quad

%%T-SuperInvk
\inferrule[(T-SuperInvk)]{
\Gamma(\this) <: \C \\\\
\forall i\in 1..n\ \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
%\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
}{\Gamma \vdash \C\QM.\QM{super}\QM.\m\QM(\overline{\e}\QM) \in \C'}

%%T-Var
\\[5ex]
\inferrule[(T-Var)]{
\Gamma(\x)=\C
}{
\Gamma \vdash \x \in\C}
\quad\quad\quad

%%T-Obj
\inferrule[(T-Obj)]{
\forall i\in 1..n\
%\Gamma_i
\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
\vdash\e_i\in \_\subtype\C^{\mh_i}\\\\
\forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
\forall i\in 1..k\ \Gamma(\x_i)\subtype\T_i\\\\
\forall\m\mbox{ such that }
\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
%\forall i\in 1\ldots n\ \Gamma_i=\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
}{
\Gamma \vdash\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM=\x_1\QM;\ldots\T_k\ \f_k\QM=\x_k\QM;\
\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
\ldots
\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
\cC
\in\C
}
\\[5ex]

%%T-update
\quad
\inferrule[(T-update)]{
\Gamma \vdash \e\in\_<:\Gamma(\x)\\\\
\Gamma \vdash \e'\in\C
}{
\Gamma \vdash \x\QM=\e\QM;\e'\in\C }
\quad\quad\quad\quad\quad

%%T-Intf
 \inferrule[(T-Intf)]{
IT(\C) = \ann\ \QM{interface}\ \C\ \QM{extends}\ \overline{I} \
\oC\methods\ \cC\\\\
 \forall \QM{default}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
\ \ \Gamma^{\mh},\,\QM{this}{:}\C\vdash\e\in \_\subtype\C^{\mh} \\\\
 \forall \QM{static}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
\ \ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
\dom(\C)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)
 }{
\C \text{ OK}
}
\end{array}$
\caption{CJ Typing}
\label{ET}
\end{figure}
\subsection{Grammar and typing rules}${}_{}$\\
